%% 
%% This is file `grid1.tex',
%% generated with the docstrip utility.
%% 
%% The original source files were:
%% 
%% crosswrd.dtx  (with options: `grid1')
%% Copyright C 1989 Brian Hamilton Kelly 
%% 
%% crossword macros for LaTeX 
%% updated for LaTeX2e by Frank Mittelbach 96/12/06 
%% 
%% for documentation run latex crosswrd.dtx 
%% 
%% distribution consist of 
%% 
%% crosswrd.dtx source 
%% crosswrd.ins installation script 
%% 00readme 
%% 
%% don't distribute this file without distributing the above 
%% files as well 
%% 
\clue{8}{A}{1}{2}{SWAM}{Points a thousand tested for witchcraft.}{4}
\clue{9}{A}{6}{2}{OHO}{Gourmet's triumphant cry on finding middle-cut
                       Pacific salmon!}{3}
\clue{10}{A}{10}{2}{ICECAP}{One hundred stride backwards across a
                            Pole.}{3-3}
\clue{11}{A}{1}{4}{BOPEEP}{Fifties' jazz record about Eastern childs'
                           play.}{2-4}
\clue{12}{A}{8}{4}{SCHEDULE}{Timetable created by editor in synagogue
                             of Spain.}{8}
\clue{13}{A}{1}{6}{THALASSOGRAPHER}{The dialect a girl mixed up tangle
                                    around symbolic diagram used by
                                    maritime student!}{15}
\clue{15}{A}{1}{8}{HAIRPIN}{Wire fastening bent road narrowly.}{7}
\clue{17}{A}{9}{8}{UMBRAGE}{Hammerhead consumes German company
                            and casts a shade.}{7}
\clue{20}{A}{1}{10}{SCALENETRIANGLE}{Strange cel alien chops to make a
                                     figure with odd sides.}{7,8}
\clue{23}{A}{1}{12}{AMOUNTED}{Sounds like a hoarse editor came to in
                              total!}{8}
\clue{25}{A}{10}{12}{ALLEGE}{Assert without proof everyone, for
                             example, English.}{6}
\clue{26}{A}{1}{14}{FLORAL}{Flourished examination of flowers.}{6}
\clue{27}{A}{8}{14}{NIL}{Floor covering discards fuming sulphuric acid
                         and returns to nothing.}{3}
\clue{28}{A}{12}{14}{TACE}{``Latin for a candle'' to be silent note
                           about aircraftman?}{4}
\clue{1}{D}{2}{1}{SWOOSH}{Water rush noise disturbs show so!}{6}
\clue{2}{D}{4}{1}{IMPELLER}{Mischievous child with cloth measure
                            hesitates to assemble rotor.}{8}
\clue{3}{D}{6}{1}{COMPASSIONATELY}{Mercifully inclined to pass round
                                   ten? Nay, about short blower!}{15}
\clue{4}{D}{8}{1}{TORSION}{Sort ion? An isotron gives it a new
                           twist!}{7}
\clue{5}{D}{10}{1}{DITHYRAMBICALLY}{Wildly and boisterously rearrange
                                    Billy May Third, roughly.}{15}
\clue{6}{D}{12}{1}{SENDUP}{Satirical book or film---give odds about
                           revision of ``Dune''?}{4-2}
\clue{7 {\noexpand\rm\&} 24}{D}{14}{1}{PALL}{Premier took in a Lord
                                    Lieutenant and all played an
                                    old game in London street.}{4-4}
\clue{14}{D}{14}{6}{ERG}{Work expended in power games?}{3}
\clue{16}{D}{2}{8}{ARC}{A church circle (or part of one).}{3}
\clue{18}{D}{12}{8}{RINGLETS}{Encircle hindrances under hair in long
                              curls.}{8}
\clue{19}{D}{8}{9}{STUDENT}{Boss over oto\-rhino\-laryng\-o\-logy
                            department undergraduate.}{7}
\clue{21}{D}{4}{10}{LOUVRE}{Old dovecote in Parisian museum.}{6}
\clue{22}{D}{14}{10}{LEGACY}{Like ornamental fabric, for example,
                             that's in bequest.}{6}
\clue{24}{D}{2}{12}{MALL}{}{See {\bf 7}}

\endinput
%% 
%% End of file `grid1.tex'.
